PANCHANGE_data_preprocessing

### 2020-08-14 PANCHANGE_data_preprocessing2020-07-15.Rmd
Finished adapting dataset combination and difference score chunks. Started adapting respiratory sections.

glad data frame list
```{r glad data frame list}
glad.dfs.list <- list(
  dem.raw.id,
  prepan.gad.raw.id,
  mhd.raw.categories,
  prepan.ocir.raw.id,
  prepan.pcl.raw.id,
  prepan.phq.raw.id,
  coping.gad.raw.id,
  coping.pcl.raw.id,
  coping.phq.raw.id,
  coping.ocir.raw.id
)
```

```{r Join all questionnaires for glad}
dat.glad.raw <- plyr::join_all(
  glad.dfs.list,
  by = c("ID", "Sample") # Alternatively you can join by several columns EDGI -> "Age", "Sex", "Birthyear"
  )

#look at the data
skimr::skim(dat.glad.raw)
```

Check duplicates again to be saved
+++KLP: RAMP: More than one column: Age, Sex, Login ID; if they match on all 3 then we treat them as a duplicate; otherwise we are going to rename the Login ID
+++KT: No duplicates identified by ID in GLAD. 
```{r glad Check duplicates}
dat.glad.raw$ID.dup <- duplicated(dat.glad.raw$ID)

summary(as.factor(dat.glad.raw$ID.dup))

dat.dup <- dat.glad.raw %>%
  filter(ID.dup == TRUE)

dat.glad.raw.no.dup <- dat.glad.raw %>%
  filter(ID.dup == FALSE)
```

#Exclusion
Exclude those without sex, gender, age and incomplete PHQ/GAD items
Create a new new object for each exclusion step, e.g., data.no_sex, data.no_gender and data.no_sex.no_age
```{r Exclude participants that do not report sex}
dim(dat.glad.raw)
dat.no.sex <- dat.glad.raw %>%
  drop_na(Sex)
dim(dat.no.sex)

dim(dat.glad.raw)[1]-dim(dat.no.sex)[1]
```
`r dim(dat.glad.raw)[1]-dim(dat.no.sex)[1]` were excluded because they did not report their sex

@@

Exclusion of inflammatory bowel disease (IBD) BioResource (BR) cohort
```{r COVID exclusion of IBD}
#dim(dat.no.sex.no.age)
#
#dat.no.sex.no.age <- dat.no.sex.no.age %>%
#  filter(NBR.IBD == 0)
#
#dim(dat.no.sex.no.age.no.ibd)
#
#dim(dat.no.sex.no.age)[1]-dim(dat.no.sex.no.age.no.ibd)[1]

dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age
```

##Create tidy data on which the exclusion criteria have been applied
Complete sex, gender, age and missingness information applied. 
```{r Tidy data}
dat.glad <- dat.no.sex.no.gender.no.age.no.ibd

skimr::skim(dat.glad)
```

@@

```{r Prepandemic definitions january_31}
dat.glad <- dat.glad %>%
  mutate(
    prepandemic_january_31_numeric =
      if_else(
        startDate.prepandemic > date_first_case_uk_january_31, # The GLAD sign up is dated before the cut off (> longer ago)
        true = 0, # Participant has not signed up before the pandemic,
        false = 1, # Participant has signed up before the pandemic,
        missing = NA_real_
      )
  )

dat.glad <- dat.glad %>%
  mutate(
    prepandemic_january_31 =
      recode_factor(
        prepandemic_january_31_numeric,
        "0" = "Sign up after 31. January",
        "1" = "Sign up before 31. January"
      )
  )

dat.glad %>%
  freq(prepandemic_january_31)
```

```{r Prepandemic definitions march_1}
dat.glad <- dat.glad %>%
  mutate(
    prepandemic_march_1_numeric =
      if_else(
        startDate.prepandemic > date_higher_awareness_march_1, # The GLAD sign up is dated before the cut off (> longer ago)
        true = 0, # Participant has not signed up before the pandemic,
        false = 1, # Participant has signed up before the pandemic,
        missing = NA_real_
      )
  )

dat.glad <- dat.glad %>%
  mutate(
    prepandemic_march_1 =
      recode_factor(
        prepandemic_march_1_numeric,
        "0" = "Sign up after 1. March",
        "1" = "Sign up before 1. March"
      )
  )

dat.glad %>%
  freq(prepandemic_march_1)
```

```{r Prepandemic definitions march_23}
dat.glad <- dat.glad %>%
  mutate(
    prepandemic_march_23_numeric =
      if_else(
        startDate.prepandemic > date_uk_lockdown_march_23, # The GLAD sign up is dated before the cut off (> longer ago)
        true = 0, # Participant has not signed up before the pandemic,
        false = 1, # Participant has signed up before the pandemic,
        missing = NA_real_
      )
  )

dat.glad <- dat.glad %>%
  mutate(
    prepandemic_march_23 =
      recode_factor(
        prepandemic_march_23_numeric,
        "0" = "Sign up after 23. March",
        "1" = "Sign up before 23. March"
      )
  )

dat.glad %>%
  freq(prepandemic_march_23)
```

Calculate the time difference between time difference between sign up questionnaire and COPING questionnaire
```{r time difference between sign up questionnaire and COPING questionnaire}
dat.glad <- dat.glad %>%
  mutate(
    time_diff_sign_up_coping =
        as.numeric(
        difftime(           #this function does time1 - time2 as difftime(time1, time2...)
          startDate.coping,    #should be start date for COPING (time1)
          startDate.prepandemic),    #should be start date for GLAD (time2)
          units = c("days")  #can also be "secs", "mins", "hours", "weeks"
          )
        )

dat.glad %>%
  descr(
    var = time_diff_sign_up_coping,
    stats = "common"
    )
```

Proportion of people signed up over one year ago
```{r eval=FALSE, include=FALSE}
dat.glad <- dat.glad %>%
  mutate(
    over_one_year_time_diff_sign_up_coping_numeric =
      as.numeric(
        if_else(time_diff_sign_up_coping > 365, 1, 0
                )
      )
  )

dat.glad <- dat.glad %>%
  mutate(
    over_one_year_time_diff_sign_up_coping =
      recode_factor(
        over_one_year_time_diff_sign_up_coping_numeric,
        "0" = "Sign up under one year before COPING",
        "1" = "Sign up over one year before COPING"
      )
  )

dat.glad %>%
  freq(over_one_year_time_diff_sign_up_coping, cumul = F)
```

Proportion of people signed up over one year ago
```{r eval=FALSE, include=FALSE}
dat.glad <- dat.glad %>%
  mutate(
    over_half_a_year_time_diff_sign_up_coping_numeric =
      as.numeric(
        if_else(time_diff_sign_up_coping > 182.5, 1, 0
                )
      )
  )

dat.glad <- dat.glad %>%
  mutate(
    over_half_a_year_time_diff_sign_up_coping =
      recode_factor(
        over_half_a_year_time_diff_sign_up_coping_numeric,
        "0" = "Sign up under half a year before COPING",
        "1" = "Sign up over half a year before COPING"
      )
  )

dat.glad %>%
  freq(over_half_a_year_time_diff_sign_up_coping, cumul = F)
```

#Total score comparison between PHQ, PCL and GAD, prepandemic, retrospecive and baseline
PHQ difference scores
```{r phq difference scores}
# Difference score for prepandemic and retrospective phq sum scores
dat.glad <- dat.glad %>%
  mutate(
    phq.diff_score_prepan_retro =
      as.numeric(
        phq.sum_score_prepan - phq.sum_score_retro  # retrospective sum score MINUS prepandemic sum score 
      )
  )

summary(dat.glad$phq.diff_score_prepan_retro)

# Difference score for prepandemic and baseline phq sum scores
dat.glad <- dat.glad %>%
  mutate(
    phq.diff_score_prepan_base =
      as.numeric(
        phq.sum_score_base - phq.sum_score_prepan  # baseline sum score MINUS prepandemic sum score 
      )
  )

summary(dat.glad$phq.diff_score_prepan_base)

# Difference score for retrospective and baseline phq sum scores
dat.glad <- dat.glad %>%
  mutate(
    phq.diff_score_base_retro =
      as.numeric(
        phq.sum_score_base - phq.sum_score_retro  # baseline sum score MINUS retrospective sum score 
      )
  )

summary(dat.glad$phq.diff_score_base_retro)

@@

GAD difference scores
```{r gad difference scores}
# Difference score for prepandemic and retrospective gad sum scores
dat.glad <- dat.glad %>%
  mutate(
    gad.diff_score_prepan_retro =
      as.numeric(
        gad.sum_score_prepan - gad.sum_score_retro  # retrospective sum score MINUS prepandemic sum score 
      )
  )

summary(dat.glad$gad.diff_score_prepan_retro)

# Difference score for prepandemic and baseline gad sum scores
dat.glad <- dat.glad %>%
  mutate(
    gad.diff_score_prepan_base =
      as.numeric(
        gad.sum_score_base - gad.sum_score_prepan  # baseline sum score MINUS prepandemic sum score 
      )
  )

summary(dat.glad$gad.diff_score_prepan_base)

# Difference score for retrospective and baseline gad sum scores
dat.glad <- dat.glad %>%
  mutate(
    gad.diff_score_base_retro =
      as.numeric(
        gad.sum_score_base - gad.sum_score_retro  # baseline sum score MINUS retrospective sum score 
      )
  )

summary(dat.glad$gad.diff_score_base_retro)

@@

PCL difference scores
NOTE: Only computed baseline and prepandemic differences as the retro account here isn't actually retrospective, but whether the reported baseline scores are due to the pandemic. 
```{r pcl difference scores}
# Difference score for prepandemic and baseline pcl sum scores
dat.glad <- dat.glad %>%
  mutate(
    pcl.diff_score_prepan_base =
      as.numeric(
        pcl.sum_score_base - pcl.sum_score_prepan  # baseline sum score MINUS prepandemic sum score 
      )
  )

summary(dat.glad$pcl.diff_score_prepan_base)
```

```{r frequency of those who reported "due to pandemic" in retro PCL}
summary(dat.glad$pcl.sum_score_retro)

freq(dat.glad$pcl.stressful_experience_repeated_images_retro)
freq(dat.glad$pcl.stressful_situation_avoiding_activities_retro)
freq(dat.glad$pcl.feeling_irritable_or_having_angry_outbursts_retro)
freq(dat.glad$pcl.stressful_experience_upset_reminded_retro)
freq(dat.glad$pcl.cut_people_feeling_distant_retro)
freq(dat.glad$pcl.difficulty_concentrating_retro)
```

OCI-R difference scores
```{r ocir difference scores}
# Difference score for prepandemic and retrospective oci-r sum scores
dat.glad <- dat.glad %>%
  mutate(
    ocir.diff_score_prepan_retro =
      as.numeric(
        ocir.sum_score_prepan - ocir.sum_score_retro  # retrospective sum score MINUS prepandemic sum score 
      )
  )

summary(dat.glad$ocir.diff_score_prepan_retro)

# Difference score for prepandemic and baseline ocir sum scores
dat.glad <- dat.glad %>%
  mutate(
    ocir.diff_score_prepan_base =
      as.numeric(
        ocir.sum_score_prepan - ocir.sum_score_base  # prepandemic sum score MINUS baseline sum score 
      )
  )

summary(dat.glad$ocir.diff_score_prepan_base)

# Difference score for retrospective and baseline ocir sum scores
dat.glad <- dat.glad %>%
  mutate(
    ocir.diff_score_base_retro =
      as.numeric(
        ocir.sum_score_base - ocir.sum_score_retro  # baseline sum score MINUS retrospective sum score 
      )
  )

summary(dat.glad$ocir.diff_score_base_retro)

@@

#Data level groups
Identify which data (prepan, retro and baseline) each participant has.
```{r data level groups PHQ}
dat.glad <- dat.glad %>%
  mutate(
    data_group_phq_numeric =
      case_when(
        na_per_person_phq_base == 0 & na_per_person_phq_prepan == 0 & na_per_person_phq_retro == 0 ~ 5, #all complete baseline, prepan and retro
        na_per_person_phq_base != 0 & na_per_person_phq_prepan != 0 & na_per_person_phq_retro == 0  ~ 4, #retro data only
        na_per_person_phq_base == 0 & na_per_person_phq_prepan != 0 & na_per_person_phq_retro != 0  ~ 3, #baseline data only
        na_per_person_phq_base == 0 & na_per_person_phq_prepan == 0 & na_per_person_phq_retro != 0  ~ 2, #prepan and baseline data only
        na_per_person_phq_base == 0 & na_per_person_phq_prepan != 0 & na_per_person_phq_retro == 0 ~ 1,  #baseline and retro data only
        na_per_person_phq_base != 0 & na_per_person_phq_prepan == 0 & na_per_person_phq_retro != 0  ~ 0  #prepan data only
      )
  )

dat.glad <- dat.glad %>%
  mutate(
    data_group_phq = 
      recode_factor(
        data_group_phq_numeric,
        "0" = "Only prepandemic data",
        "1" = "Baseline and retrospective data",
        "2" = "Prepandemic and baseline data",
        "3" = "Baseline data only",
        "4" = "Retrospective data only",
        "5" = "All prepandemic, baseline and retrospective data"
      )
  )

table(dat.glad$data_group_phq)
```

GAD
```{r data level groups GAD}
dat.glad <- dat.glad %>%
  mutate(
    data_group_gad_numeric =
      case_when(
        na_per_person_gad_base == 0 & na_per_person_gad_prepan == 0 & na_per_person_gad_retro == 0 ~ 5, #all complete baseline, prepan and retro
        na_per_person_gad_base != 0 & na_per_person_gad_prepan != 0 & na_per_person_gad_retro == 0  ~ 4, #retro data only
        na_per_person_gad_base == 0 & na_per_person_gad_prepan != 0 & na_per_person_gad_retro != 0  ~ 3, #baseline data only
        na_per_person_gad_base == 0 & na_per_person_gad_prepan == 0 & na_per_person_gad_retro != 0  ~ 2, #prepan and baseline data only
        na_per_person_gad_base == 0 & na_per_person_gad_prepan != 0 & na_per_person_gad_retro == 0 ~ 1,  #baseline and retro data only
        na_per_person_gad_base != 0 & na_per_person_gad_prepan == 0 & na_per_person_gad_retro != 0  ~ 0  #prepan data only
      )
  )

dat.glad <- dat.glad %>%
  mutate(
    data_group_gad = 
      recode_factor(
        data_group_gad_numeric,
        "0" = "Only prepandemic data",
        "1" = "Baseline and retrospective data",
        "2" = "Prepandemic and baseline data",
        "3" = "Baseline data only",
        "4" = "Retrospective data only",
        "5" = "All prepandemic, baseline and retrospective data"
      )
  )

table(dat.glad$data_group_gad)
```

PCL
```{r data level groups PCL}
dat.glad <- dat.glad %>%
  mutate(
    data_group_pcl_numeric =
      case_when(
        na_per_person_pcl_base == 0 & na_per_person_pcl_prepan == 0 & na_per_person_pcl_retro == 0 ~ 5, #all complete baseline, prepan and retro
        na_per_person_pcl_base != 0 & na_per_person_pcl_prepan != 0 & na_per_person_pcl_retro == 0  ~ 4, #retro data only
        na_per_person_pcl_base == 0 & na_per_person_pcl_prepan != 0 & na_per_person_pcl_retro != 0  ~ 3, #baseline data only
        na_per_person_pcl_base == 0 & na_per_person_pcl_prepan == 0 & na_per_person_pcl_retro != 0  ~ 2, #prepan and baseline data only
        na_per_person_pcl_base == 0 & na_per_person_pcl_prepan != 0 & na_per_person_pcl_retro == 0 ~ 1,  #baseline and retro data only
        na_per_person_pcl_base != 0 & na_per_person_pcl_prepan == 0 & na_per_person_pcl_retro != 0  ~ 0  #prepan data only
      )
  )

dat.glad <- dat.glad %>%
  mutate(
    data_group_pcl = 
      recode_factor(
        data_group_pcl_numeric,
        "0" = "Only prepandemic data",
        "1" = "Baseline and retrospective data",
        "2" = "Prepandemic and baseline data",
        "3" = "Baseline data only",
        "4" = "Retrospective data only",
        "5" = "All prepandemic, baseline and retrospective data"
      )
  )

table(dat.glad$data_group_pcl)
```

#Read in respiratory data
```{r coping resp cols}
coping.glad.respiratory.raw <- readRDS(file = paste0(data.raw_path, "/glad/respiratory_coping_glad.rds"))
dim(coping.glad.respiratory.raw)
colnames(coping.glad.respiratory.raw)
```

>>>>

glad/edgi data frame list
```{r glad/edgi data frame list}
if(GLAD == TRUE | EDGI == TRUE) {
dfs.list <- list(
  dem.raw.id,
  prepan.gad.raw.id,
  mhd.raw.categories,
  prepan.ocir.raw.id,
  prepan.pcl.raw.id,
  prepan.phq.raw.id,
  coping.gad.raw.id,
  coping.pcl.raw.id,
  coping.phq.raw.id,
  coping.ocir.raw.id
)
}
```

Removing all prepandemic dataframes from NBR/RAMP (prepandemic measures don't exist for these cohorts)
```{r nbr/ramp data frame list}
if(NBR == TRUE | RAMP == TRUE) {
dfs.list <- list(
  dem.raw.id,
  mhd.raw.categories,
  coping.gad.raw.id,
  coping.pcl.raw.id,
  coping.phq.raw.id,
  coping.ocir.raw.id
)
}
```

```{r Join all questionnaires}
dat.raw <- plyr::join_all(
  dfs.list,
  by = c("ID", "Sample") # Alternatively you can join by several columns EDGI -> "Age", "Sex", "Birthyear"
  )

#look at the data
skimr::skim(dat.raw)
```

Check duplicates again to be saved
+++KLP: RAMP: More than one column: Age, Sex, Login ID; if they match on all 3 then we treat them as a duplicate; otherwise we are going to rename the Login ID
+++KT: No duplicates identified by ID in GLAD. 
+++MD: Duplicates have been removed in cleaning script for NBR. No duplicates at this stage.
```{r Check duplicates}
dat.raw$ID.dup <- duplicated(dat.raw$ID)

summary(as.factor(dat.raw$ID.dup))

dat.dup <- dat.raw %>%
  filter(ID.dup == TRUE)

dat.raw.no.dup <- dat.raw %>%
  filter(ID.dup == FALSE)
```

#Exclusion
Exclude those without sex, gender, age and incomplete PHQ/GAD items
Create a new new object for each exclusion step, e.g., data.no_sex, data.no_gender and data.no_sex.no_age
```{r Exclude participants that do not report sex}
dim(dat.raw)
dat.no.sex <- dat.raw %>%
  drop_na(Sex)
dim(dat.no.sex)

dim(dat.raw)[1]-dim(dat.no.sex)[1]
```
`r dim(dat.glad.raw)[1]-dim(dat.no.sex)[1]` were excluded because they did not report their sex

@@

Exclusion of inflammatory bowel disease (IBD) BioResource (BR) cohort
++MD: 3,410 participants from NBR IBD cohort excluded
```{r COVID exclusion of IBD}
if(NBR == TRUE) {
dim(dat.no.sex.no.gender.no.age)

dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age %>%
  filter(Cohort != "IBD")

dim(dat.no.sex.no.gender.no.age.no.ibd)
dim(dat.no.sex.no.gender.no.age)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd)[1]
}

if(GLAD == TRUE | EDGI == TRUE | RAMP == TRUE) {
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age
}
```

##Create tidy data on which the exclusion criteria have been applied
Complete sex, gender, age and missingness information applied. 
```{r Tidy data}
dat <- dat.no.sex.no.gender.no.age.no.ibd

skimr::skim(dat)
```

@@

```{r Prepandemic definitions january_31}
if(GLAD == TRUE | EDGI == TRUE){
dat <- dat %>%
  mutate(
    prepandemic_january_31_numeric =
      if_else(
        startDate.prepandemic > date_first_case_uk_january_31, # The GLAD sign up is dated before the cut off (> longer ago)
        true = 0, # Participant has not signed up before the pandemic,
        false = 1, # Participant has signed up before the pandemic,
        missing = NA_real_
      )
  )

dat <- dat %>%
  mutate(
    prepandemic_january_31 =
      recode_factor(
        prepandemic_january_31_numeric,
        "0" = "Sign up after 31. January",
        "1" = "Sign up before 31. January"
      )
  )

dat %>%
  freq(prepandemic_january_31)
}
```

```{r Prepandemic definitions march_1}
if(GLAD == TRUE | EDGI == TRUE){
dat <- dat %>%
  mutate(
    prepandemic_march_1_numeric =
      if_else(
        startDate.prepandemic > date_higher_awareness_march_1, # The GLAD sign up is dated before the cut off (> longer ago)
        true = 0, # Participant has not signed up before the pandemic,
        false = 1, # Participant has signed up before the pandemic,
        missing = NA_real_
      )
  )

dat <- dat %>%
  mutate(
    prepandemic_march_1 =
      recode_factor(
        prepandemic_march_1_numeric,
        "0" = "Sign up after 1. March",
        "1" = "Sign up before 1. March"
      )
  )

dat %>%
  freq(prepandemic_march_1)
}
```

```{r Prepandemic definitions march_23}
if(GLAD == TRUE | EDGI == TRUE){
dat <- dat %>%
  mutate(
    prepandemic_march_23_numeric =
      if_else(
        startDate.prepandemic > date_uk_lockdown_march_23, # The GLAD sign up is dated before the cut off (> longer ago)
        true = 0, # Participant has not signed up before the pandemic,
        false = 1, # Participant has signed up before the pandemic,
        missing = NA_real_
      )
  )

dat <- dat %>%
  mutate(
    prepandemic_march_23 =
      recode_factor(
        prepandemic_march_23_numeric,
        "0" = "Sign up after 23. March",
        "1" = "Sign up before 23. March"
      )
  )

dat %>%
  freq(prepandemic_march_23)
}
```

Calculate the time difference between time difference between sign up questionnaire and COPING questionnaire
```{r time difference between sign up questionnaire and COPING questionnaire}
if(GLAD == TRUE | EDGI == TRUE){
dat <- dat %>%
  mutate(
    time_diff_sign_up_coping =
        as.numeric(
        difftime(           #this function does time1 - time2 as difftime(time1, time2...)
          startDate.coping,    #should be start date for COPING (time1)
          startDate.prepandemic),    #should be start date for GLAD (time2)
          units = c("days")  #can also be "secs", "mins", "hours", "weeks"
          )
        )

dat %>%
  descr(
    var = time_diff_sign_up_coping,
    stats = "common"
    )
}
```

Proportion of people signed up over one year ago
```{r glad/edgi ppts signed up over one year ago, eval=FALSE, include=FALSE}
if(GLAD == TRUE | EDGI == TRUE){
dat <- dat %>%
  mutate(
    over_one_year_time_diff_sign_up_coping_numeric =
      as.numeric(
        if_else(time_diff_sign_up_coping > 365, 1, 0
                )
      )
  )

dat <- dat %>%
  mutate(
    over_one_year_time_diff_sign_up_coping =
      recode_factor(
        over_one_year_time_diff_sign_up_coping_numeric,
        "0" = "Sign up under one year before COPING",
        "1" = "Sign up over one year before COPING"
      )
  )

dat %>%
  freq(over_one_year_time_diff_sign_up_coping, cumul = F)
}
```

Proportion of people signed up over half a year ago
```{r glad/edgi ppts signed up over half a year ago, r eval=FALSE, include=FALSE}
if(GLAD == TRUE | EDGI == TRUE){
dat <- dat %>%
  mutate(
    over_half_a_year_time_diff_sign_up_coping_numeric =
      as.numeric(
        if_else(time_diff_sign_up_coping > 182.5, 1, 0
                )
      )
  )

dat <- dat %>%
  mutate(
    over_half_a_year_time_diff_sign_up_coping =
      recode_factor(
        over_half_a_year_time_diff_sign_up_coping_numeric,
        "0" = "Sign up under half a year before COPING",
        "1" = "Sign up over half a year before COPING"
      )
  )

dat %>%
  freq(over_half_a_year_time_diff_sign_up_coping, cumul = F)
}
```

#Total score comparison between PHQ, PCL and GAD, prepandemic, retrospecive and baseline
PHQ difference scores
```{r phq difference scores}
if(GLAD == TRUE | EDGI == TRUE){
# Difference score for prepandemic and retrospective phq sum scores
dat <- dat %>%
  mutate(
    phq.diff_score_prepan_retro =
      as.numeric(
        phq.sum_score_prepan - phq.sum_score_retro  # retrospective sum score MINUS prepandemic sum score 
      )
  )

summary(dat$phq.diff_score_prepan_retro)

# Difference score for prepandemic and baseline phq sum scores
dat <- dat %>%
  mutate(
    phq.diff_score_prepan_base =
      as.numeric(
        phq.sum_score_base - phq.sum_score_prepan  # baseline sum score MINUS prepandemic sum score 
      )
  )

summary(dat$phq.diff_score_prepan_base)
}
# Difference score for retrospective and baseline phq sum scores
dat <- dat %>%
  mutate(
    phq.diff_score_base_retro =
      as.numeric(
        phq.sum_score_base - phq.sum_score_retro  # baseline sum score MINUS retrospective sum score 
      )
  )

summary(dat$phq.diff_score_base_retro)

@@

GAD difference scores
```{r gad difference scores}
if(GLAD == TRUE | EDGI == TRUE){
# Difference score for prepandemic and retrospective gad sum scores
dat <- dat %>%
  mutate(
    gad.diff_score_prepan_retro =
      as.numeric(
        gad.sum_score_prepan - gad.sum_score_retro  # retrospective sum score MINUS prepandemic sum score 
      )
  )

summary(dat$gad.diff_score_prepan_retro)

# Difference score for prepandemic and baseline gad sum scores
dat <- dat %>%
  mutate(
    gad.diff_score_prepan_base =
      as.numeric(
        gad.sum_score_base - gad.sum_score_prepan  # baseline sum score MINUS prepandemic sum score 
      )
  )

summary(dat$gad.diff_score_prepan_base)
}
# Difference score for retrospective and baseline gad sum scores
dat <- dat %>%
  mutate(
    gad.diff_score_base_retro =
      as.numeric(
        gad.sum_score_base - gad.sum_score_retro  # baseline sum score MINUS retrospective sum score 
      )
  )

summary(dat$gad.diff_score_base_retro)

@@

PCL difference scores
NOTE: Only computed baseline and prepandemic differences as the retro account here isn't actually retrospective, but whether the reported baseline scores are due to the pandemic. 
```{r pcl difference scores}
if(GLAD == TRUE | EDGI == TRUE){
# Difference score for prepandemic and baseline pcl sum scores
dat <- dat %>%
  mutate(
    pcl.diff_score_prepan_base =
      as.numeric(
        pcl.sum_score_base - pcl.sum_score_prepan  # baseline sum score MINUS prepandemic sum score 
      )
  )

summary(dat$pcl.diff_score_prepan_base)
}
```

```{r frequency of those who reported "due to pandemic" in retro PCL}
summary(dat$pcl.sum_score_retro)

freq(dat$pcl.stressful_experience_repeated_images_retro)
freq(dat$pcl.stressful_situation_avoiding_activities_retro)
freq(dat$pcl.feeling_irritable_or_having_angry_outbursts_retro)
freq(dat$pcl.stressful_experience_upset_reminded_retro)
freq(dat$pcl.cut_people_feeling_distant_retro)
freq(dat$pcl.difficulty_concentrating_retro)
```

OCI-R difference scores
```{r ocir difference scores}
if(GLAD == TRUE | EDGI == TRUE){
# Difference score for prepandemic and retrospective oci-r sum scores
dat <- dat %>%
  mutate(
    ocir.diff_score_prepan_retro =
      as.numeric(
        ocir.sum_score_prepan - ocir.sum_score_retro  # retrospective sum score MINUS prepandemic sum score 
      )
  )

summary(dat$ocir.diff_score_prepan_retro)

# Difference score for prepandemic and baseline ocir sum scores
dat <- dat %>%
  mutate(
    ocir.diff_score_prepan_base =
      as.numeric(
        ocir.sum_score_prepan - ocir.sum_score_base  # prepandemic sum score MINUS baseline sum score 
      )
  )

summary(dat$ocir.diff_score_prepan_base)
}
# Difference score for retrospective and baseline ocir sum scores
dat <- dat %>%
  mutate(
    ocir.diff_score_base_retro =
      as.numeric(
        ocir.sum_score_base - ocir.sum_score_retro  # baseline sum score MINUS retrospective sum score 
      )
  )

summary(dat$ocir.diff_score_base_retro)

@@

#Data level groups
Identify which data (prepan, retro and baseline) each participant has.
```{r glad/edgi data level groups PHQ}
if(GLAD == TRUE | EDGI == TRUE){
dat <- dat %>%
  mutate(
    data_group_phq_numeric =
      case_when(
        na_per_person_phq_base == 0 & na_per_person_phq_prepan == 0 & na_per_person_phq_retro == 0 ~ 5, #all complete baseline, prepan and retro
        na_per_person_phq_base != 0 & na_per_person_phq_prepan != 0 & na_per_person_phq_retro == 0  ~ 4, #retro data only
        na_per_person_phq_base == 0 & na_per_person_phq_prepan != 0 & na_per_person_phq_retro != 0  ~ 3, #baseline data only
        na_per_person_phq_base == 0 & na_per_person_phq_prepan == 0 & na_per_person_phq_retro != 0  ~ 2, #prepan and baseline data only
        na_per_person_phq_base == 0 & na_per_person_phq_prepan != 0 & na_per_person_phq_retro == 0 ~ 1,  #baseline and retro data only
        na_per_person_phq_base != 0 & na_per_person_phq_prepan == 0 & na_per_person_phq_retro != 0  ~ 0  #prepan data only
      )
  )

dat <- dat %>%
  mutate(
    data_group_phq = 
      recode_factor(
        data_group_phq_numeric,
        "0" = "Only prepandemic data",
        "1" = "Baseline and retrospective data",
        "2" = "Prepandemic and baseline data",
        "3" = "Baseline data only",
        "4" = "Retrospective data only",
        "5" = "All prepandemic, baseline and retrospective data"
      )
  )

table(dat$data_group_phq)
}
```

```{r nbr/ramp data level groups PHQ}
if(NBR == TRUE | RAMP == TRUE){
dat <- dat %>%
  mutate(
    data_group_phq_numeric =
      case_when(
#        na_per_person_phq_base == 0 & na_per_person_phq_prepan == 0 & na_per_person_phq_retro == 0 ~ 5, #all complete baseline, prepan and retro
        na_per_person_phq_base != 0 & na_per_person_phq_retro == 0  ~ 4, #retro data only
        na_per_person_phq_base == 0 & na_per_person_phq_retro != 0  ~ 3, #baseline data only
#        na_per_person_phq_base == 0 & na_per_person_phq_prepan == 0 & na_per_person_phq_retro != 0  ~ 2, #prepan and baseline data only
        na_per_person_phq_base == 0 & na_per_person_phq_retro == 0 ~ 1,  #baseline and retro data only
#        na_per_person_phq_base != 0 & na_per_person_phq_prepan == 0 & na_per_person_phq_retro != 0  ~ 0  #prepan data only
      )
  )

dat <- dat %>%
  mutate(
    data_group_phq = 
      recode_factor(
        data_group_phq_numeric,
        "0" = "Only prepandemic data",
        "1" = "Baseline and retrospective data",
        "2" = "Prepandemic and baseline data",
        "3" = "Baseline data only",
        "4" = "Retrospective data only",
        "5" = "All prepandemic, baseline and retrospective data"
      )
  )

table(dat$data_group_phq)
}
```

GAD
```{r glad/edgi data level groups GAD}
if(GLAD == TRUE | EDGI == TRUE){
dat <- dat %>%
  mutate(
    data_group_gad_numeric =
      case_when(
        na_per_person_gad_base == 0 & na_per_person_gad_prepan == 0 & na_per_person_gad_retro == 0 ~ 5, #all complete baseline, prepan and retro
        na_per_person_gad_base != 0 & na_per_person_gad_prepan != 0 & na_per_person_gad_retro == 0  ~ 4, #retro data only
        na_per_person_gad_base == 0 & na_per_person_gad_prepan != 0 & na_per_person_gad_retro != 0  ~ 3, #baseline data only
        na_per_person_gad_base == 0 & na_per_person_gad_prepan == 0 & na_per_person_gad_retro != 0  ~ 2, #prepan and baseline data only
        na_per_person_gad_base == 0 & na_per_person_gad_prepan != 0 & na_per_person_gad_retro == 0 ~ 1,  #baseline and retro data only
        na_per_person_gad_base != 0 & na_per_person_gad_prepan == 0 & na_per_person_gad_retro != 0  ~ 0  #prepan data only
      )
  )

dat <- dat %>%
  mutate(
    data_group_gad = 
      recode_factor(
        data_group_gad_numeric,
        "0" = "Only prepandemic data",
        "1" = "Baseline and retrospective data",
        "2" = "Prepandemic and baseline data",
        "3" = "Baseline data only",
        "4" = "Retrospective data only",
        "5" = "All prepandemic, baseline and retrospective data"
      )
  )

table(dat$data_group_gad)
}
```

```{r nbr/ramp data level groups GAD}
if(NBR == TRUE | RAMP == TRUE){
dat <- dat %>%
  mutate(
    data_group_gad_numeric =
      case_when(
#        na_per_person_gad_base == 0 & na_per_person_gad_prepan == 0 & na_per_person_gad_retro == 0 ~ 5, #all complete baseline, prepan and retro
        na_per_person_gad_base != 0 & na_per_person_gad_retro == 0  ~ 4, #retro data only
        na_per_person_gad_base == 0 & na_per_person_gad_retro != 0  ~ 3, #baseline data only
#        na_per_person_gad_base == 0 & na_per_person_gad_prepan == 0 & na_per_person_gad_retro != 0  ~ 2, #prepan and baseline data only
        na_per_person_gad_base == 0 & na_per_person_gad_retro == 0 ~ 1,  #baseline and retro data only
#        na_per_person_gad_base != 0 & na_per_person_gad_prepan == 0 & na_per_person_gad_retro != 0  ~ 0  #prepan data only
      )
  )

dat <- dat %>%
  mutate(
    data_group_gad = 
      recode_factor(
        data_group_gad_numeric,
        "0" = "Only prepandemic data",
        "1" = "Baseline and retrospective data",
        "2" = "Prepandemic and baseline data",
        "3" = "Baseline data only",
        "4" = "Retrospective data only",
        "5" = "All prepandemic, baseline and retrospective data"
      )
  )

table(dat$data_group_gad)
}
```

PCL
```{r glad/edgi data level groups PCL}
if(GLAD == TRUE | EDGI == TRUE){
dat <- dat %>%
  mutate(
    data_group_pcl_numeric =
      case_when(
        na_per_person_pcl_base == 0 & na_per_person_pcl_prepan == 0 & na_per_person_pcl_retro == 0 ~ 5, #all complete baseline, prepan and retro
        na_per_person_pcl_base != 0 & na_per_person_pcl_prepan != 0 & na_per_person_pcl_retro == 0  ~ 4, #retro data only
        na_per_person_pcl_base == 0 & na_per_person_pcl_prepan != 0 & na_per_person_pcl_retro != 0  ~ 3, #baseline data only
        na_per_person_pcl_base == 0 & na_per_person_pcl_prepan == 0 & na_per_person_pcl_retro != 0  ~ 2, #prepan and baseline data only
        na_per_person_pcl_base == 0 & na_per_person_pcl_prepan != 0 & na_per_person_pcl_retro == 0 ~ 1,  #baseline and retro data only
        na_per_person_pcl_base != 0 & na_per_person_pcl_prepan == 0 & na_per_person_pcl_retro != 0  ~ 0  #prepan data only
      )
  )

dat <- dat %>%
  mutate(
    data_group_pcl = 
      recode_factor(
        data_group_pcl_numeric,
        "0" = "Only prepandemic data",
        "1" = "Baseline and retrospective data",
        "2" = "Prepandemic and baseline data",
        "3" = "Baseline data only",
        "4" = "Retrospective data only",
        "5" = "All prepandemic, baseline and retrospective data"
      )
  )

table(dat$data_group_pcl)
}
```
